---
title: "Run R CMD check from R and Capture Results"
output:
  github_document:
    toc: true
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcmdcheck

> Run R CMD check from R and Capture Results

<!-- badges: start -->
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![](https://www.r-pkg.org/badges/version/rcmdcheck)](https://www.r-pkg.org/pkg/rcmdcheck)
[![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/rcmdcheck)](https://www.r-pkg.org/pkg/rcmdcheck)
[![R-CMD-check](https://github.com/r-lib/rcmdcheck/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/r-lib/rcmdcheck/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/r-lib/rcmdcheck/branch/main/graph/badge.svg)](https://app.codecov.io/gh/r-lib/rcmdcheck?branch=main)
<!-- badges: end -->

Run R CMD check from R programmatically and capture the results of the
individual checks.

## Installation

Install the released version from CRAN
 
```r
install.packages("rcmdcheck")
```

Or install the development version from GitHub:

```r
# install.packages("pak")
pak::pak("r-lib/rcmdcheck")
```

## Usage

```r
library(rcmdcheck)
rcmdcheck("path/to/R/package")
```

Call `rcmdcheck()` on a source R package `.tar.gz` file, or on a folder
containing your R package. Supply `quiet = FALSE` if you want to omit the
output. The result of the check is returned, in a list with elements
`errors`, `warnings`, and `notes`. Each element is a character vector,
and one element of the character vectors is a single failure.

<img width="1000" src="https://cdn.jsdelivr.net/gh/r-lib/rcmdcheck@main/tools/rcmdcheck.svg" alt="animated screenshot of a terminal window demonstrating example usage of the rcmdcheck function.">

### Programmatic usage

`rcmdcheck()` returns an `rcmdcheck` object, which you can query and
manipulate.

```{r}
library(rcmdcheck)
chk <- rcmdcheck("tests/testthat/bad1", quiet = TRUE)
chk
```

`check_details()` turns the check results into a simple lists with the
following information currently:

```{r}
names(check_details(chk))
```

* `package`: Package name.
* `version`: Package version number.
* `notes`: Character vector of check `NOTE`s.
* `warnings`: Character vector of check `WARNING`s.
* `errors`: Character vector of check `ERROR`s.
* `platform`: Platform, e.g. `x86_64-apple-darwin15.6.0`.
* `checkdir`: Check directory.
* `install_out`: Output of the package installation.
* `description`: The text of the `DESCRIPTION` file.
* `session_info`: A `sessioninfo::session_info` object, session information
  from within the check process.
* `cran`: Flag, whether this is a CRAN package. (Based on the `Repository`
  field in `DESCRIPTION`, which is typically only set for published CRAN
  packages.)
* `bioc`: Flag, whether this is a Bioconductor package, based on the
  presence of the `biocViews` field in `DESCRIPTION`.

Note that if the check results were parsed from a file, some of these
fields might be missing (`NULL`), as we don't have access to the original
`DESCRIPTION`, the installation output, etc.

### Parsing check output

`parse_check()` parses check output from a file, `parse_check_url()`
parses check output from a URL.

### CRAN checks

rcmdcheck has a functions to access CRAN's package check results.

`cran_check_flavours()` downloads the names of the CRAN platforms:

```{r}
cran_check_flavours()
```

`cran_check_results()` loads and parses all check results for a package.

```{r}
cran_check_results("igraph")
```

### Comparing checks

`compare_checks()` can compare two or more `rcmdcheck` objects.
`compare_to_cran()` compares an `rcmdcheck` object to the CRAN checks of
the same package:

```{r}
chk <- rcmdcheck(quiet = TRUE)
compare_to_cran(chk)
```

### Background processes

`rcmdcheck_process` is a `processx::process` class, that can run
`R CMD check` in the background. You can also use this to run multiple
checks concurrently. `processx::process` methods can be used to poll or
manipulate the check processes.

```{r}
chkpx <- rcmdcheck_process$new()
chkpx
```

```{r}
chkpx$wait()
chkpx$parse_results()
```
